Area maximisation in continuous time

### 2013-04-25 Notes/ContTime.tex
wrote up some finite time max stuff

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
%
\begin{equation}\label{eq:rowsum}
  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
  \qquad
  \mathbf{p}(t)\onev=1,
  \qquad
  \MM\onev=0.
\end{equation}
%
In Surya's notes, $\MM=r\W^F$.

@@

The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
%
\begin{equation}\label{eq:MWdef}
  \M^\pm \equiv \I + \W^\pm,
  \qquad
  \M^\pm\onev = \onev,
  \qquad
  \M^\pm_{ij} \in [0,1].
\end{equation}
%
The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
%
\begin{equation}\label{eq:forgetting}
  \MM = r \W^F \equiv r\prn{f^+\W^+ + f^-\W^-}.
\end{equation}
%
We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.

@@

As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
%
\begin{equation}\label{eq:signal}
  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
\end{equation}
%
where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
%
\begin{equation}\label{eq:initprob}
  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\W^F}.
\end{equation}
%
We can rewrite the signal as
%
\begin{equation}\label{eq:signalr}
\begin{aligned}
  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\W^F}\w\\
    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
\end{aligned}
\end{equation}
%
For the noise, we can use the standard deviation of the synaptic strength:
%
\begin{equation}\label{eq:noise}
  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
\end{equation}
%
From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
%
\begin{equation}\label{eq:plusminusstates}
  \boldsymbol{+} = \set{k}{\w_k=+1},
  \qquad
  \boldsymbol{-} = \set{k}{\w_k=-1}.
\end{equation}
%
This simplifies the noise:
%
\begin{equation}\label{eq:noisepm}
  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
\end{equation}
%
The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
\end{equation}
%

We can express this in terms of the one parameter family of transition matrices:
%
\begin{equation}\label{eq:Walpha}
  \begin{aligned}
  \W(\alpha) &= \alpha\W^++(1-\alpha)\W^-,
  &\quad&\implies&
    \W^F&=\W(f^+),\\&&&&
    \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \W^F.
  \end{aligned}
\end{equation}
%
Then \eqref{eq:SNRcurve} becomes
%
\begin{equation}\label{eq:SNRalpha}
  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\W^F)\,\e^{rt\W^F}\w.
\end{equation}
%

@@

Suppose that we have a partition such that $\W^+$ and $\W^-$ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
%
\begin{equation}\label{eq:lumpablesynapse}
  VU\W^\pm V = \W^\pm V,
  \qquad
  VU\w=\w.
\end{equation}
%
We can define a new synapse with
%
\begin{equation}\label{eq:lumpedsynapse}
  \widehat{\W}^\pm = U\W^\pm V,
  \qquad
  \widehat{\w} = U \w,
  \qquad
  \widehat{\mathbf{p}}^\infty = \eq V.
\end{equation}
%
This synapse has an SNR curve:
%
\begin{equation}\label{eq:lumpedSNR}
  \begin{aligned}
    \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
      &= \eq VU (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\W^F}\w. \\
  \end{aligned}
\end{equation}
%
\ie the lumped process has exactly the same SNR as the original one.

@@

The flux will be maximised if potentiation is guaranteed to take it to a $+$ state and depression is guaranteed to take it to a $-$ state:
%
\begin{equation}\label{eq:guarantee}
  \sum_{j\in+} \W^+_{ij} = 1 \quad \forall\, i\in-,
  \qquad
  \sum_{j\in-} \W^-_{ij} = 1 \quad \forall\, i\in+.
\end{equation}
%
This implies that the process is lumpable to a two state system (see \sref{sec:lump} and \sref{sec:SNRlump}):
%
\begin{multline}\label{eq:binarylump}
%\begin{aligned}
  \W^+ = \begin{pmatrix}
           -1 & 1 \\
           0 & 0 \\
         \end{pmatrix}
  ,\quad
  \W^- = \begin{pmatrix}
           0 & 0 \\
           1 & -1 \\
         \end{pmatrix}
  ,\quad
  \w = \begin{pmatrix}
           -1 \\
           1 \\
         \end{pmatrix}
  ,\\ \implies\quad
  \W^F = \begin{pmatrix}
           -f^+ & f^+ \\
           f^- & -f^- \\
         \end{pmatrix}
  ,\quad
  \eq = \begin{pmatrix}
           f^- & f^+ \\
         \end{pmatrix}
  ,%&
%  \e^{rt\W^F} &= \I+(1-\e^{-rt})\W^F
%         \begin{pmatrix}
%           f^-+f^+\e^{-rt} & f^+(1-\e^{-rt}) \\
%           f^-(1-\e^{-rt}) & f^++f^-\e^{-rt} \\
%         \end{pmatrix}
%  ,
%  \end{aligned}
\end{multline}
%
which leads to the initial SNR %curve
\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to
%$\SNR(t) = \sqrt{N (4f^+f^-)} \,\e^{-rt}$,
$\SNR(0) \leq \sqrt{N (4f^+f^-)}$,
which would have no effect on what follows.}
%
\begin{equation}\label{eq:binarySNR}
 % \SNR(t) = \sqrt{N} (4f^+f^-) \,\e^{-rt}.
  \SNR(0) \leq \sqrt{N} (4f^+f^-) .
\end{equation}
%
%The initial SNR is maximised at $f^+=f^-=\half$:
This is maximised at $f^+=f^-=\half$:
%
\begin{equation}\label{eq:initmax}
  \SNR(0) \leq \sqrt{N}.
\end{equation}
%

@@

The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
%%
%\begin{equation}\label{eq:SNR}
%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\W^F) \e^{rt\W^F} \w.
%\end{equation}
%%
The area is computed by integrating this
%
\begin{equation}\label{eq:area}
\begin{aligned}
  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\W^F}}_0^\infty \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \w.
\end{aligned}
\end{equation}
%

We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $q_{ij} \equiv \diff{\W^F_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
%
\begin{equation}\label{eq:areaT}
  \hat{A} = \sum_{i,j,k} \eq_i q_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
\end{equation}
%

@@

We can rewrite \eqref{eq:areaT} as
%
\begin{equation}\label{eq:areaEta}
\begin{aligned}
  \hat{A} &= 2\sum_{i,j} q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
    &= -2\sum_{i,j} q_{ij} \eq_i \eta^+_{j} \\
    &= 2\sum_{i,j} q_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
    &= 2\sum_{i,j} q_{ij} \eq_i \eta^-_{j}.
\end{aligned}
\end{equation}
%
We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
%
\begin{equation}\label{eq:areaZ}
  \hat{A} = \sum_{i,j,k,l} q_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
    = \pib \fund q \fund \w.
\end{equation}
%

It is also helpful to define the following quantities:
%
\begin{equation}\label{eq:areacoeffs}
  \begin{aligned}
    c_k &= \diff{\ln\eq_k}{\alpha}
      = \sum_{ij} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
      = - \prn{\eq q \fptb}_k
      = \frac{(\eq q \fund)_k}{\eq_k}, \\
    a_i &= \sum_j q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
    \implies
    \hat{A} &= \sum_k c_k \eq_k \w_k
      = 2\sum_i a_i.
  \end{aligned}
\end{equation}
%
Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.

@@

As discussed in \sref{sec:constraints},
we will regard the off-diagonal elements of $\W^\pm_{ij}$ to be the independent variables,
with $\W^\pm_{ii}=-\sum_{j \neq i} \W^\pm_{ij}$ imposed by hand.
Thus,
%
\begin{equation}\label{eq:basicderivs}
  \pdiff{\W^F_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
  \qquad
  \pdiff{q_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
\end{equation}
%
The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.

In particular, differentiating \eqref{eq:funddef},
%
\begin{equation}\label{eq:derivZ}
  \pdiff{\fund_{ij}}{\W^\pm_{gh}} = rf^\pm \fund_{ig} (\fund_{hj}-\fund_{gj}).
\end{equation}
%
We can then differentiate expression \eqref{eq:areaZ} to get
%
\begin{equation}\label{eq:derivA}
  \begin{aligned}
    \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
      rf^\pm \sum_{ijkl} q_{ij} \w_k \pib_l \brk{
         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
    =&\,
      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g \brk{
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
    =&\,
      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f^\pm)}} \\
    =&\,
      2rf^\pm \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
  \end{aligned}
\end{equation}
%
where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
%
\begin{equation}\label{eq:rowsum}
  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
  \qquad
  \mathbf{p}(t)\onev=1,
  \qquad
  \MM\onev=0.
\end{equation}
%
In Surya's notes, $\MM=r\frg$.

@@

The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
%
\begin{equation}\label{eq:MWdef}
  \M^\pm \equiv \I + \W^\pm,
  \qquad
  \M^\pm\onev = \onev,
  \qquad
  \M^\pm_{ij} \in [0,1].
\end{equation}
%
The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
%
\begin{equation}\label{eq:forgetting}
  \MM = r \frg \equiv r\prn{f^+\W^+ + f^-\W^-}.
\end{equation}
%
We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.

@@

As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
%
\begin{equation}\label{eq:signal}
  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
\end{equation}
%
where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
%
\begin{equation}\label{eq:initprob}
  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\frg}.
\end{equation}
%
We can rewrite the signal as
%
\begin{equation}\label{eq:signalr}
\begin{aligned}
  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\frg}\w\\
    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
\end{aligned}
\end{equation}
%
For the noise, we can use the standard deviation of the synaptic strength:
%
\begin{equation}\label{eq:noise}
  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
\end{equation}
%
From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
%
\begin{equation}\label{eq:plusminusstates}
  \boldsymbol{+} = \set{k}{\w_k=+1},
  \qquad
  \boldsymbol{-} = \set{k}{\w_k=-1}.
\end{equation}
%
This simplifies the noise:
%
\begin{equation}\label{eq:noisepm}
  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
\end{equation}
%
The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
\end{equation}
%

We can express this in terms of the one parameter family of transition matrices:
%
\begin{equation}\label{eq:Walpha}
  \begin{aligned}
  \W(\alpha) &= \alpha\W^++(1-\alpha)\W^-,
  &\quad&\implies&
    \frg&=\W(f^+),\\&&&&
    \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
  \end{aligned}
\end{equation}
%
Then \eqref{eq:SNRcurve} becomes
%
\begin{equation}\label{eq:SNRalpha}
  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
\end{equation}
%

@@

Suppose that we have a partition such that $\W^+$ and $\W^-$ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
%
\begin{equation}\label{eq:lumpablesynapse}
  VU\W^\pm V = \W^\pm V,
  \qquad
  VU\w=\w.
\end{equation}
%
We can define a new synapse with
%
\begin{equation}\label{eq:lumpedsynapse}
  \widehat{\W}^\pm = U\W^\pm V,
  \qquad
  \widehat{\w} = U \w,
  \qquad
  \widehat{\mathbf{p}}^\infty = \eq V.
\end{equation}
%
This synapse has an SNR curve:
%
\begin{equation}\label{eq:lumpedSNR}
  \begin{aligned}
    \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
      &= \eq VU (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
      &= \eq (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\frg} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\frg}\w. \\
  \end{aligned}
\end{equation}
%
\ie the lumped process has exactly the same SNR as the original one.

@@

The flux will be maximised if potentiation is guaranteed to take it to a $+$ state and depression is guaranteed to take it to a $-$ state:
%
\begin{equation}\label{eq:guarantee}
  \sum_{j\in+} \W^+_{ij} = 1 \quad \forall\, i\in-,
  \qquad
  \sum_{j\in-} \W^-_{ij} = 1 \quad \forall\, i\in+.
\end{equation}
%
This implies that the process is lumpable to a two state system (see \sref{sec:lump} and \sref{sec:SNRlump}):
%
\begin{multline}\label{eq:binarylump}
%\begin{aligned}
  \W^+ = \begin{pmatrix}
           -1 & 1 \\
           0 & 0 \\
         \end{pmatrix}
  ,\quad
  \W^- = \begin{pmatrix}
           0 & 0 \\
           1 & -1 \\
         \end{pmatrix}
  ,\quad
  \w = \begin{pmatrix}
           -1 \\
           1 \\
         \end{pmatrix}
  ,\\ \implies\quad
  \frg = \begin{pmatrix}
           -f^+ & f^+ \\
           f^- & -f^- \\
         \end{pmatrix}
  ,\quad
  \eq = \begin{pmatrix}
           f^- & f^+ \\
         \end{pmatrix}
  ,%&
%  \e^{rt\frg} &= \I+(1-\e^{-rt})\frg
%         \begin{pmatrix}
%           f^-+f^+\e^{-rt} & f^+(1-\e^{-rt}) \\
%           f^-(1-\e^{-rt}) & f^++f^-\e^{-rt} \\
%         \end{pmatrix}
%  ,
%  \end{aligned}
\end{multline}
%
which leads to the initial SNR %curve
\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to
%$\SNR(t) = \sqrt{N (4f^+f^-)} \,\e^{-rt}$,
$\SNR(0) \leq \sqrt{N (4f^+f^-)}$,
which would have no effect on what follows.}
%
\begin{equation}\label{eq:binarySNR}
 % \SNR(t) = \sqrt{N} (4f^+f^-) \,\e^{-rt}.
  \SNR(0) \leq \sqrt{N} (4f^+f^-) .
\end{equation}
%
%The initial SNR is maximised at $f^+=f^-=\half$:
This is maximised at $f^+=f^-=\half$:
%
\begin{equation}\label{eq:initmax}
  \SNR(0) \leq \sqrt{N}.
\end{equation}
%

@@

The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
%%
%\begin{equation}\label{eq:SNR}
%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
%\end{equation}
%%
The area is computed by integrating this
%
\begin{equation}\label{eq:area}
\begin{aligned}
  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \w.
\end{aligned}
\end{equation}
%

We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
%
\begin{equation}\label{eq:areaT}
  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
\end{equation}
%

@@

We can rewrite \eqref{eq:areaT} as
%
\begin{equation}\label{eq:areaEta}
\begin{aligned}
  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
\end{aligned}
\end{equation}
%
We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
%
\begin{equation}\label{eq:areaZ}
  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
    = \pib \fund q \fund \w.
\end{equation}
%

It is also helpful to define the following quantities:
%
\begin{equation}\label{eq:areacoeffs}
  \begin{aligned}
    c_k &= \diff{\ln\eq_k}{\alpha}
      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
      = - \prn{\eq q \fptb}_k
      = \frac{(\eq q \fund)_k}{\eq_k}, \\
    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
    \implies
    \hat{A} &= \sum_k c_k \eq_k \w_k
      = 2\sum_i a_i.
  \end{aligned}
\end{equation}
%
Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.

@@

As discussed in \sref{sec:constraints},
we will regard the off-diagonal elements of $\W^\pm_{ij}$ to be the independent variables,
with $\W^\pm_{ii}=-\sum_{j \neq i} \W^\pm_{ij}$ imposed by hand.
Thus,
%
\begin{equation}\label{eq:basicderivs}
  \pdiff{\frg_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
  \qquad
  \pdiff{\enc_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
\end{equation}
%
The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.

In particular, differentiating \eqref{eq:funddef},
%
\begin{equation}\label{eq:derivZ}
  \pdiff{\fund_{ij}}{\W^\pm_{gh}} = rf^\pm \fund_{ig} (\fund_{hj}-\fund_{gj}).
\end{equation}
%
We can then differentiate expression \eqref{eq:areaZ} to get
%
\begin{equation}\label{eq:derivA}
  \begin{aligned}
    \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
      rf^\pm \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
    =&\,
      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
    =&\,
      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f^\pm)}} \\
    =&\,
      2rf^\pm \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
  \end{aligned}
\end{equation}
%
where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.

It is sometimes useful to consider the following derivatives:
%
\begin{equation}\label{eq:derivqw}
  \begin{aligned}
    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
    \pdiff{}{\enc_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
  \end{aligned}
\end{equation}
%
Each of these derivatives behaves as their names suggest:
%
\begin{equation}\label{eq:derivqweff}
  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
  \qquad
  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
\end{equation}
%
This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.
